---
title: "Homework 8"
output: html_document
date: "2025-03-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading data

```{r}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation

# My data
z <- read.csv("/Users/kristofrm/Downloads/obesity_data.csv")
colnames(z)[colnames(z) == "Weight"] <- "myVar"
str(z)
summary(z)
```

``` {r, warning=FALSE}
# Plotting histogram
real_data <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
# Add empirical density curve
real_data <-  real_data +  geom_density(linetype="dotted",size=0.75)

# Plotting
real_data
```

``` {r}
# Get parameters for normal
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute

# Plot normal probability density
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
real_data + stat
```

``` {r}
# Plot exponential probability density
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
real_data + stat + stat2
```

``` {r}
# Plot uniform probability density
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
real_data + stat + stat2 + stat3
```

``` {r}
# Plot gamma probability density
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

gamma_distribution_line <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
real_data + gamma_distribution_line
```

``` {r}
# Plot beta probability density
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

## Simulated data based on gamma distribution parameters
```{r}
# Getting parameters
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]
length_data <- gammaPars$n

# Making gamma distribution using above parameters
simulated_gamma <- rgamma(n=length_data,shape=shapeML,rate=rateML)
```

``` {r}
# Plotting histogram
df <- data.frame(x=simulated_gamma)
estimated_data <- ggplot(data=df, aes(x=x, y=..density..)) +
  geom_histogram(color="black",fill="goldenrod",size=0.2) 

# Getting distribution line
stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))

estimated_data + stat4
```

``` {r}
# Real data again
real_data + gamma_distribution_line
```

#### How do the two histogram profiles compare? Do you think the model is doing a good job of simulating realistic data that match your original measurements? Why or why not?

The two histogram profiles are pretty close. The gamma distribution does a decent job modeling the actual data. However, the histogram based on the gamma distribution parameters does not match the actual data histogram too closely. For the most part, however, the model is doing a decent job simualting realistic data that matches the original measurements. If there was an option to make the histogram data fit the distribution line a bit less perfectly, it would be ideal.







